---
title: "Test Vanilla Model"
output:
  html_document: default
  html_notebook: default
---
In this notebook, we show the performance of the vanilla model with 3 layers (where sensitivity datsets have been combined into a single matrix to increase overlap). This is essentially the new benchmark to beat in terms of auc.

In order not to lose our sanity with the relative file paths that rmd chooses to employ, let's set the root directory of this project

```{r setup}
knitr::opts_knit$set(root.dir = normalizePath('..'))
getwd()
```


We start by importing the necessary functions

```{r}
rm(list=ls())

source("RCode/flexible_layers/structureDataFlexible.R")
source("RCode/flexible_layers/sensitivityDataFlexible.R")
source("RCode/flexible_layers/perturbationDataFlexible.R")
source("RCode/flexible_layers/luminexDataFlexible.R")
source("RCode/flexible_layers/imagingDataFlexible.R")

source("RCode/flexible_layers/constSensitivityLayerFlexible.R")
source("RCode/flexible_layers/constStructureLayerFlexible.R")
source("RCode/flexible_layers/constLuminexLayerFlexible.R")
source("RCode/flexible_layers/constImagingLayerFlexible.R")
source("RCode/flexible_layers/constPerturbationLayerFlexible.R")

source("RCode/flexible_layers/integrateLayersFlexible.R")
source("RCode/flexible_layers/generateDrugPairsFlexible.R")
source("RCode/flexible_layers/drugTargetBenchFlexible.R")
source("RCode/flexible_layers/compConcordIndxFlexible.R")
source("RCode/flexible_layers/printCIndices.R")
source("RCode/flexible_layers/printPVals.R")
source("RCode/flexible_layers/generateROCPlotFlexible.R")
source("RCode/flexible_layers/atcBenchFlexible.R")
source("RCode/flexible_layers/compDrugTargetBenchmarkFlexible.R")
source("RCode/flexible_layers/computeATCBenchmarkFlexible.R")
source("RCode/flexible_layers/stackedLayerAnalysisHelpers.R")
source("RCode/flexible_layers/communityGenFlexible.R")

source("RCode/cindexComp2.R")
source("RCode/predPerf.R")

library(PharmacoGx)
library(apcluster)
library(rcdk)
library(fingerprint)
library(annotate)
library(org.Hs.eg.db)
library(SNFtool)
library(ROCR)
library(survcomp)
library(reshape2)
library(proxy)
library(PRROC)
library(apcluster)

# Load lincs metadata and clean the pert_inames in it
badchars <- "[\xb5]|[\n]|[,]|[;]|[:]|[-]|[+]|[*]|[%]|[$]|[#]|[{]|[}]|[[]|[]]|[|]|[\\^]|[/]|[\\]|[.]|[_]|[ ]"
lincs.meta <- read.csv("Data/LINCS.csv", stringsAsFactors = FALSE)
lincs.meta$pert_iname <- toupper(lincs.meta$pert_iname)
lincs.meta$pert_iname <- gsub(badchars, "", lincs.meta$pert_iname)

pert.file.name <- "Data/L1000_compound_signatures.RData"
sensitivity.file.name <- "Data/combined_sensitivity//combined_sens_iname_replaced.RData"
source("RCode/flexible_layers/stackedLayerAnalysis.R")
```


We begin with evaluating the performance of our model on the ctrpv2 dataset.
```{r}
res <- EvaluateModelROC(use.sensitivity=TRUE, use.perturbation=TRUE, use.structure=TRUE,
            use.imaging=FALSE, use.luminex=FALSE, 
             sensitivity.file.name=sensitivity.file.name, pert.file.name=pert.file.name, 
            lincs.meta=lincs.meta,
             atc.benchmark.name="chembl-new", compute.atc=FALSE, use.ctrpv2=TRUE, use.clue=FALSE,
             use.chembl=FALSE, use.dbank=FALSE, use.dtc=FALSE, create.communities=FALSE, base.dir="Output/auc_p_flex")
```