---
title: "Verify AUC Correctness"
output: html_notebook
---

```{r setup}
knitr::opts_knit$set(root.dir = normalizePath("../../"))
```


```{r}
source("RCode/custom_c_index_computation/computeCIndex.R")
library(ROCR)
```

We first look at a linearly seperable case

```{r}
x <- c(0.1, 0.3, 0.6, 0.8, 0.9)
y <- c(0, 0, 1, 1, 1)

custom.c.index <- ComputeCIndex(x, y)
pred <- prediction(x, as.factor(y))
pert <- performance(pred, 'tpr', 'fpr')
```


Now we see performance on random data

```{r}
x <- rnorm(100, 0.55, 0.05)
y <- x + rnorm(100, -0.1, 0.05)
y <- round(y)

custom.c.index <- ComputeCIndex(x, y)
pred <- prediction(x, as.factor(y))
pert <- performance(pred, 'tpr', 'fpr')
auc <-  performance(pred, 'auc')
```

Lastly, let's look at random linearly separable data

```{r}
x <- rnorm(100, 0.55, 0.05)
y <- round(x)

custom.c.index <- ComputeCIndex(x, y)
pred <- prediction(x, as.factor(y))
pert <- performance(pred, 'tpr', 'fpr')
auc <-  performance(pred, 'auc')
```



Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).
